---
title: "Summary Statistics and Figures"
author: "Chase Webber"
subtitle: Summary Statistics and Figures
output:
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

library(knitr)

# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = FALSE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```


<!-- ### start answering your problem set here -->
<!-- You may export your homework in either html or pdf, with the former usually being easier. 
     To export or compile your Rmd file: click above on 'Knit' then 'Knit to HTML' -->
<!-- Be sure to submit both your .Rmd file and the compiled .html or .pdf file for full credit -->


```{r setup-3}

# load all your libraries in this chunk 

library("tidyverse")
library("partykit")


library("PerformanceAnalytics")
library("rpart")
library("rpart.plot")


library("randomForest")

library("rsample")
library("randomForestExplainer")

library("DALEX")
library("pdp")

library("visNetwork")
library("sparkline")

library("ggplot2")
library("ggridges")
# note, do not run install.packages() inside a code chunk. install them in the console outside of a code chunk. 

```



## Cleaned and Transformed dataset

```{r}

matches <- read.csv("archive3/WorldCupMatches.csv")
players <- read.csv("archive3/WorldCupPlayers.csv")
world_cups <- read.csv("archive3/WorldCups.csv")
shoot_outs <- read.csv("archive2/shootouts.csv")

matches_clean <- matches %>% 
  mutate(total_goals = (matches$Home.Team.Goals + matches$Away.Team.Goals))

matches_clean <- matches_clean %>% 
  mutate(team_win = (ifelse(Home.Team.Goals>Away.Team.Goals, 'Home', ifelse(Home.Team.Goals<Away.Team.Goals, 'Away', 'Draw')))) %>% drop_na()

home_df <- data.frame(Country = matches$Home.Team.Name, Goals = matches$Home.Team.Goals, Year = matches$Year)
away_df <- data.frame(Country = matches$Away.Team.Name, Goals = matches$Away.Team.Goals, Year = matches$Year)

home <- na.omit(home_df)
away <- na.omit(away_df)

country_goals <- rbind(home, away)

shoot_outs <- 
  shoot_outs %>% 
  mutate(winner = 
           as_factor(winner))

players_split <- initial_split(players, prop = 0.75) 
cups_split <- initial_split(world_cups, prop = 0.75) 
fifa_split <- initial_split(matches_clean, prop = 0.75) 

players_train <- training(players_split)
players_test <- testing(players_split)

cups_train <- training(cups_split)
cups_test <- testing(cups_split)

fifa_train <- training(fifa_split)
fifa_test <- testing(fifa_split)

```


## Summary Statistics

```{r}

country_goals %>% arrange(desc(Goals)) %>% select(Country,Goals,Year) %>% head(n=10)

table(world_cups$Winner)

mean(country_goals$Goals)

max(world_cups$GoalsScored)

min(world_cups$GoalsScored)

```
Summary Statistics include:
- Top 10 most amount of goals scored by a country in a World Cup match
- Table of winners of the World Cup and how many times they have won
- Average amount of goals scored per team in a World Cup match: 1.42 goals
- Most amount of goals scored in a World Cup: 171
- Least amount of goals scored in a World Cup: 70

## Total Amount of Goals Scored in World Cup

```{r}

ggplot(data = world_cups, aes(x = Year, y = GoalsScored)) + geom_col() + labs(y = "Goals Scored", caption = "*World Cup was not held in 1942 and 1946", title = "Total Goals Scored per year")


```

Chart showing how total goals scored have increased over time. 

view(matches_clean)

## Top 50 Countries in Total Goals Scored in World Cup

```{r fig.height=10}

goals_country <- country_goals %>% select(-Year)
goals_country <- country_goals %>% mutate(Country = as.factor(Country))


goals_hist <- goals_country %>% ggplot() + aes(x = Goals, y = fct_lump_n(fct_rev(fct_infreq(Country)), n = 20), label=Goals) + geom_col() + labs(x = "Total Goals Scored in World Cup", y = "Country", title = "Top 20 Countries in World Cup Goals Scored", caption = "*disregard other")+theme(axis.text=element_text(size=30),
        axis.title=element_text(size=30,face="bold"))
goals_hist 


ggplot(goals_country, aes(fct_lump_n(fct_rev(fct_infreq(Country)), n = 20))) + geom_histogram(stat = 'count', binwidth = 50) + coord_flip() + labs(x = "Country", y = "Total Games Played in World Cup", title = "Top 20 Countries in World Cup Games Played", caption = "*disregard other")+theme(axis.text=element_text(size=30),
        axis.title=element_text(size=30,face="bold"))






```

Chart showing the amount of goals scored all-time per country in the world cup. 

## Home Field Advantage

```{r}

matches_clean <- matches_clean %>% mutate(team_win = as.factor(team_win))
country_goals <- country_goals %>% mutate(country_simple = fct_lump(Country, n = 10))
matches_clean <- matches_clean %>% mutate(home_simple = fct_lump(Home.Team.Name, n = 10))
matches_clean <- matches_clean %>% mutate(away_simple = fct_lump(Away.Team.Name, n = 10))


home_chart <- ggplot(data = matches_clean, aes(x="", y = team_win, fill = team_win))+
  geom_bar(width=1, stat = "identity")

home_pie <- home_chart + coord_polar("y", start = 0) + labs(title = "Match Outcome per Home or Away", y = "Outcome", x = "Team")
home_pie

```

Pie chart showing win share between home and away teams.

## Top 50 Countries in Total Shootout Wins in World Cup

```{r}

shoot_outs <- shoot_outs %>% mutate(winner_topten = fct_lump_n(winner, n = 10))

shoot_outs <- shoot_outs %>% mutate(winner_topten = fct_infreq(winner_topten))

winner_subset <- shoot_outs[shoot_outs$winner_topten %in% c('South Korea', 
                                                          'Zambia',
                                                          'Egypt',
                                                          'Argentina',
                                                          'South Africa',
                                                          'Thailand',
                                                          'Kenya ',
                                                          'Brazil',
                                                          'Iraq',
                                                          'Senegal'),]
ggplot(winner_subset, aes(x = fct_rev(winner_topten))) + 
  geom_histogram(stat = "count") +
  coord_flip() + labs(title = "Top 10 Shootout Winners", x = "Country", y = "Shootout wins")

```

Chart showing top 10 countries with the most shootout wins in World Cup history


